---
title: "Experiment3: More airline data"
subtitle: "Asynchronous lecture"
date:  "2023-09-10"
output: html_notebook
---

This experiment (conceptually) continues our class work on flight data.

We will continue to refer to (https://r4ds.hadley.nz/data-transform).

**However, we will expand and update our dataset by going to a root source --- the Bureau of Transportation.


 
Go to website  https://www.transtats.bts.gov/DL_SelectFields.aspx?gnoyr_VQ=FIM&QO_fu146_anzr=Nv4%20Pn44vr45 .  This dataset allows access to information on domestic flight and domestic carriers, with monthly summary data.



# Packages


```{r}
library(nycflights13)
library(tidyverse)
library(janitor)
library(here)
#> ── Attaching core tidyverse packages ───────────────────── tidyverse 2.0.0 ──
#> ✔ dplyr     1.1.2     ✔ readr     2.1.4
#> ✔ forcats   1.0.0     ✔ stringr   1.5.0
#> ✔ ggplot2   3.4.3     ✔ tibble    3.2.1
#> ✔ lubridate 1.9.2     ✔ tidyr     1.3.0
#> ✔ purrr     1.0.2     
#> ── Conflicts ─────────────────────────────────────── tidyverse_conflicts() ──
#> ✖ dplyr::filter() masks stats::filter()
#> ✖ dplyr::lag()    masks stats::lag()
#> ℹ Use the conflicted package (<http://conflicted.r-lib.org/>) to force all conflicts to become errors
```

# Get data from the online source.


Save the download to your data_raw directory. It will be a zipfile, but that will be OK.  Don't change the name of the zipfile.


## Load the BOT data

We take advantage of the `here` package to get the path to the file that we stored.  We also may note that inside the zipfile is a csv file.  We will read the csvfile.



```{r}
thisfile=here("data_raw","DL_SelectFields.zip")

df2022=read_csv(thisfile) %>% clean_names()
```

## Subsetting to data of interest

Let's focus on flights from La Guardia (airport "LGA") and eliminate cargo flights by requiring at leat 1 passenger.

Call the resultant dataframe `df9`.

```{r}
df9=df2022 |> 
  filter(passengers>0,origin=="LGA")

```


# Exploring the BOT 2022 data

## Subsetting to data of interest

Lets try to visualize a monthly count of flights using a barchart.


First, a not quite right visualization.

```{r}
df9 |> 
  ggplot(aes(month)) + geom_bar()
```
By default, `geom_bar` is counting the number of rows, where we have asked it to visualize the count by `month`.  

** Take a look at the dataset and discover why counting rows is not going to give us a count of flights. **

The visualization we want is:

```{r}
df9 |> ggplot(aes(month)) + geom_bar(aes(weight=departures_performed))
```
Make some observations about this plot.


### A new visualization

Can you make a boxplot where the bar height is based on the number of passengers riding each month.

```{r}
df9 |> ggplot(aes(month)) + geom_bar(aes(weight=passengers))
```

Observation:
* low passenger numbers in early 2022 likely due to covid.


### Just because you can


Here is pretty, but not so useful plot.

```{r}
df9 |> 
  ggplot(aes(month,fill=carrier_name)) + geom_bar(aes(weight=departures_performed))
```
## Arrivals and departures from LGA

```{r}
df10=df2022 |> filter(passengers>0,origin=="LGA" | dest=="LGA")

df10 |> ggplot(aes(month)) + geom_bar(aes(weight=passengers))
```

```{r}
df11=df10 |> 
  select(month,passengers,seats, carrier_name, dest, origin)
```

```{r}
df12=df10 |> select(1:5, month, contains("id") )

```
 
```{r}
df13=df11 |> mutate(percent_loading = passengers/seats*100)

df13 |> 
  ggplot(aes(percent_loading)) +
  geom_histogram()+facet_wrap(~carrier_name,scales="free_y")

```


### Summarize data

```{r}
df14=df2022 %>% filter(passengers>0,origin=="LGA" | dest=="LGA") %>% 
  group_by(month,carrier_name) %>% 
  summarise(total_pass=sum(passengers),  num_flights=sum(departures_performed)) %>%
  mutate(average_pass_per_flight=total_pass/num_flights)


df14 
```



