---
title: "Experiment001: MA132 enrollment"
subtitle: "In class work"
date:  "2023-10-02"
output: html_notebook
---


We will continue to refer to (https://r4ds.hadley.nz/data-transform).

Our dataset is peoplesoft querry which gives enrollment in all math sections over several semesters. 

Our goal is to predict Spring 2024 enrollment for MA132.



# Packages


```{r}

library(tidyverse)
library(janitor)
library(here)
#> ── Attaching core tidyverse packages ───────────────────── tidyverse 2.0.0 ──
#> ✔ dplyr     1.1.2     ✔ readr     2.1.4
#> ✔ forcats   1.0.0     ✔ stringr   1.5.0
#> ✔ ggplot2   3.4.3     ✔ tibble    3.2.1
#> ✔ lubridate 1.9.2     ✔ tidyr     1.3.0
#> ✔ purrr     1.0.2     
#> ── Conflicts ─────────────────────────────────────── tidyverse_conflicts() ──
#> ✖ dplyr::filter() masks stats::filter()
#> ✖ dplyr::lag()    masks stats::lag()
#> ℹ Use the conflicted package (<http://conflicted.r-lib.org/>) to force all conflicts to become errors
```

# Get data from the online source.

Completed on moodle and loaded into `data_raw` directory.

```{r}
df1=read_csv(here("data_raw","CU_ACAD_DEPT_WHEN_OFF_ENRL.csv")) %>% clean_names()

```

```{r}
skimr::skim(df1)
```

# Some initial subsetting of the data

I will remove variables that I know are irrelevant and filter to only the rows of interest (MA131 or MA132).

```{r}
df_calc=df1 |>
  select(-acad_org,-title,-current_status,-course_id) |> 
  filter(subject=="MA",
         catalog %in% c(120,125,131,132),
         component=="LEC")
```

