---
title: "Experiment1: NYC airline data"
date:  "2023-09-06"
output: html_notebook
---

This notebook is our first experiment in class.  We will be (loosely) following the analysis from R4DS chapter 4 (https://r4ds.hadley.nz/data-transform).


# Packages

Data is from `nycflights13`.  The other packages will support our analysis.  I am including `here` so that we have an easy way to reference file locations.

```{r}
library(nycflights13)
library(tidyverse)
library(janitor)
library(here)
#> ── Attaching core tidyverse packages ───────────────────── tidyverse 2.0.0 ──
#> ✔ dplyr     1.1.2     ✔ readr     2.1.4
#> ✔ forcats   1.0.0     ✔ stringr   1.5.0
#> ✔ ggplot2   3.4.3     ✔ tibble    3.2.1
#> ✔ lubridate 1.9.2     ✔ tidyr     1.3.0
#> ✔ purrr     1.0.2     
#> ── Conflicts ─────────────────────────────────────── tidyverse_conflicts() ──
#> ✖ dplyr::filter() masks stats::filter()
#> ✖ dplyr::lag()    masks stats::lag()
#> ℹ Use the conflicted package (<http://conflicted.r-lib.org/>) to force all conflicts to become errors
```

# Loading the dataset

Following from the book, we load the data, but I will save it as variable `df1`.

```{r}
df1=flights
```

And we briefly examine:

```{r}
glimpse(df1)
```
```{r}
head(df1)
```

# Other datasets to create

Our first tasking was to create datasets as follows:
* Task 1:    df2 =   flights from miami  in september
* Task2:  df3 = flights in September going to Miami
* Task3:   df4=flights in January going to Miami
* Task4:  df5=flights in summer going to Chicago

## Task 1

df2 =   flights from miami  in september

```{r}
df2=df1 |> filter(origin=="MIA",month==9)
```

It looks like this dataset only includes flights originating in New York area.


## Task 2

df3 = flights in September going to Miami

```{r}
df3=df1 |> filter(dest=="MIA",month==9)
```

We find 912 flights.

## Task 3

Task3:   df4=flights in January going to Miami

```{r}
df4=df1 |> filter(dest=="MIA",month==1)
```

There are 981 January flights to Miami.

##  Task4:  

df5=flights in summer going to Chicago


I had to look up on line to find airports in the Chicago area.

On website https://theskydeck.com/chicago-airports/ I found:

* Chicago Midway Airport Code: MDW
* Chicago O’Hare Airport Code: ORD

For summer, I will use months June, July, and August.

```{r}
df5=df1 |> 
  filter(dest %in%  c("MDW","ORD"),month %in% c(6,7,8))
```

Perhaps a better solution to the "in summer question" is to use the date information.

```{r}
df5a=df1 |> 
  filter(dest %in%  c("MDW","ORD")) |>
  filter(time_hour>="2013-06-21", time_hour<="2013-09-21")
           
```

# Day two: class tasks

* Sometimes flight numbers are repeated.  For flights to MIAMI in September, find the unique flight numbers.
* Create a dataframe df6 composed of flights to Miami with a flight number matching the smallest flight number from the set above.
* Create a visualization (using ggplot) to show the DELAYS associated with this flight number.

## Finding unique flight numbers

We start from `df3` for this question (Sep flights to Miami) and look for distinct flights, which we will sort:

```{r}
df_flight_num=df3 |>
  distinct(flight) |>
  arrange(flight)
```

Although there were 912 flights, there were only 57 flight numbers.

```{r}
head(df_flight_num)
```

## Flight 83 to Miami  
Create a dataframe df6 composed of flights to Miami with a flight number matching the smallest flight number from the set above.

```{r}
df6=df1 %>% filter(flight==83,dest=="MIA")
```

632 flights match this condition.

## Visualization of delays for flight 83

Create a visualization (using ggplot) to show the DELAYS associated with this flight number.


```{r}
df6 |> 
  ggplot(aes(x=dep_delay,y=arr_delay)) +geom_point()
```

