---
title: "Experiment2: NYC airline data"
date:  "2023-09-10"
output: html_notebook
---

This experiment continues are class work on the NYCflights data, with a bit more looking at delays.

We will continue to refer to (https://r4ds.hadley.nz/data-transform).


# Packages

Data is from `nycflights13`.  The other packages will support our analysis.  I am including `here` so that we have an easy way to reference file locations.

```{r}
library(nycflights13)
library(tidyverse)
library(janitor)
library(here)
#> ── Attaching core tidyverse packages ───────────────────── tidyverse 2.0.0 ──
#> ✔ dplyr     1.1.2     ✔ readr     2.1.4
#> ✔ forcats   1.0.0     ✔ stringr   1.5.0
#> ✔ ggplot2   3.4.3     ✔ tibble    3.2.1
#> ✔ lubridate 1.9.2     ✔ tidyr     1.3.0
#> ✔ purrr     1.0.2     
#> ── Conflicts ─────────────────────────────────────── tidyverse_conflicts() ──
#> ✖ dplyr::filter() masks stats::filter()
#> ✖ dplyr::lag()    masks stats::lag()
#> ℹ Use the conflicted package (<http://conflicted.r-lib.org/>) to force all conflicts to become errors
```

# Loading the dataset

Following from the book, we load the data, but I will save it as variable `df1`.

```{r}
df1=flights
```




# Day two: after class tasks


```{r}
df6=df1 %>% filter(flight==83,dest=="MIA")
```

146 flights match this condition.

## A first of delays for flight 83

Create a visualization (using ggplot) to show the DELAYS associated with this flight number.


```{r}
df6 |> 
  ggplot(aes(x=dep_delay,y=arr_delay)) +geom_point()
```

## Tasking at home: discover something more about delay

> Questions: 
> * Is flight time affected by delayed departure.  (Do the airlines try to "catch up"?)
> * Does the departure delay change across time of day (later in the day has more delays.)
> * Is flight time pattern affected by time of year
> * Is departure delay affected by time of year.

* Note to self

*Note to self*

**Note to self**

***Note to self***

## A second visualization:

```{r}
df1 |> 
  filter(dest=="MIA") |>
  count(origin,sort=TRUE)
```

### Is flight time affected by delayed departure.

I want to examine whether the flight time is impacted by delayed departure.

I want to compare flight time to *planned* flight time.   So we create a new variable

flt_delta=arr_delay-dep_delay

An flight that arrives 10 minute late, if it departed on time, had a "delta" of 10 minutes.


```{r}
df7=df1 |>
  filter(dest=="MIA",origin=="LGA") |>
  mutate(flt_delta=arr_delay-dep_delay)
```

```{r}
df7 |> 
  ggplot(aes(x=dep_delay,y=flt_delta)) +geom_point(alpha=.1)
```
```{r}
df7 |> 
  ggplot(aes(x=dep_delay,y=flt_delta)) +geom_point(alpha=.1)+
  geom_hline(aes(yintercept=mean(flt_delta,na.rm=TRUE)))
```

### Is departure delay affected by time of year

```{r}
df7 |>
  ggplot(aes(x=time_hour,y=dep_delay)) +geom_point(alpha=.1)
  
```

```{r}
df7 |>
  ggplot(aes(x=time_hour,y=dep_delay)) +
  geom_point(alpha=.1)+
  stat_smooth()+
  ylim(-25,120)
  
```

***Why are delays bigger in december than in january --- probably not weather.**

### Does the departure delay change across time of day?

```{r}
df7 |>
  ggplot(aes(x=hour+minute/60,y=dep_delay)) +
  geom_point(alpha=.1)+
  stat_smooth()+
  ylim(-25,120)
  
```

Observation:
* Departure delay increase across the flight day.


```{r}
df7 |> 
  mutate(day_of_week=weekdays(time_hour)) |>
  ggplot(aes(x=hour+minute/60,y=dep_delay,color=day_of_week)) +
  geom_point(alpha=.1)+
  stat_smooth()+
  ylim(-25,120)
  
```

```{r}
df7 |> 
  mutate(day_of_week=weekdays(time_hour)) |>
  ggplot(aes(x=hour+minute/60,y=dep_delay,color=day_of_week)) +
  geom_point(alpha=.1)+
  stat_smooth()+
  ylim(-20,40)+
  facet_wrap(~day_of_week)
  
```
