---
title: "Experiment5: Multiyear analysis of - red vs blue COVID impact"
subtitle: "In class work"
date:  "2023-09-20"
output: html_notebook
---


We will continue to refer to (https://r4ds.hadley.nz/data-transform).

We will use data from BOT website: 

 
Go to website  https://www.transtats.bts.gov/DL_SelectFields.aspx?gnoyr_VQ=FIM&QO_fu146_anzr=Nv4%20Pn44vr45 .  This dataset allows access to information on domestic flight and domestic carriers, with monthly summary data.

We will analyze data from 2017-present (as of Sep 2023).



# Packages


```{r}
library(nycflights13)
library(tidyverse)
library(janitor)
library(here)
#> ── Attaching core tidyverse packages ───────────────────── tidyverse 2.0.0 ──
#> ✔ dplyr     1.1.2     ✔ readr     2.1.4
#> ✔ forcats   1.0.0     ✔ stringr   1.5.0
#> ✔ ggplot2   3.4.3     ✔ tibble    3.2.1
#> ✔ lubridate 1.9.2     ✔ tidyr     1.3.0
#> ✔ purrr     1.0.2     
#> ── Conflicts ─────────────────────────────────────── tidyverse_conflicts() ──
#> ✖ dplyr::filter() masks stats::filter()
#> ✖ dplyr::lag()    masks stats::lag()
#> ℹ Use the conflicted package (<http://conflicted.r-lib.org/>) to force all conflicts to become errors
```

# Get data from the online source.

Completed on moodle and loaded into `data_raw` directory.


## Load the BOT data

We take advantage of the `here` package to get the path to the file that we stored.  We also may note that inside the zipfile is a csv file.  We will read the csvfile.



```{r}
thisfile=here("data_raw","DL_SelectFields_2017.zip")
df2017=read_csv(thisfile) %>% clean_names()

thisfile=here("data_raw","DL_SelectFields_2018.zip")
df2018=read_csv(thisfile) %>% clean_names()

thisfile=here("data_raw","DL_SelectFields_2019.zip")
df2019=read_csv(thisfile) %>% clean_names()

thisfile=here("data_raw","DL_SelectFields_2020.zip")
df2020=read_csv(thisfile) %>% clean_names()

thisfile=here("data_raw","DL_SelectFields_2021.zip")
df2021=read_csv(thisfile) %>% clean_names()

thisfile=here("data_raw","DL_SelectFields_2022.zip")
df2022=read_csv(thisfile) %>% clean_names()

thisfile=here("data_raw","DL_SelectFields_2023.zip")
df2023=read_csv(thisfile) %>% clean_names()

```

### combine dataframes 

```{r}
dfc=df2017 |>
  bind_rows(df2018) |>
  bind_rows(df2019) |>
  bind_rows(df2020) |>
  bind_rows(df2021) |>
  bind_rows(df2022) |>
  bind_rows(df2023)

```

### Create new file of combined data for easier future loading

I will save this data as csv file with intent that future loading can simply read this new file.

```{r}
write_csv(dfc,here("data","bot_flight_2017_2023m08.csv"))
```

Test readung:

```{r}
dfc1=read_csv(here("data","bot_flight_2017_2023m08.csv"))
```




## Subsetting to data of interest


### Arrivals and departures from LGA

```{r}
dfc1=dfc %>%
  filter(passengers>0,origin=="LGA" | dest=="LGA") %>%
  mutate(by_month=ym(str_c(year,"-",month))) 


# piped
# mtcars %>%
#   group_by(cyl) %>% 
#   summarise(mean_mpg = mean(mpg)) %>%
#   Unravel::unravel()

```

We visualize total LGA traffic, by month, over the analysis dataframe.

```{r}

dfc1 |> ggplot(aes(by_month)) + geom_bar(aes(weight=passengers)) +
  ggtitle("total passengers by month over multiyear period")
```

And we can visualize by year.

```{r}

dfc1 |> ggplot(aes(year)) + geom_bar(aes(weight=passengers)) +
  ggtitle("total passengers by year over multiyear period")
```

###  Question - do more folks travel *LGA to MIA* or *MIA to LGA*?

```{r}
dfc2=dfc1 |> filter(dest == "MIA" | origin=="MIA")
```

```{r}

dfc2 |> ggplot(aes(by_month,fill=origin)) + 
  geom_bar(aes(weight=passengers),position="dodge") +
  ggtitle("total passengers by month over multiyear period, between LGA and MIA")
```

### Let's use grouping to try to understand this issue better

Let's create a new dataframe, where we use "grouping" to get at values we want.

```{r}
dfc2 |>
  count(origin,by_month) |> 
  ggplot(aes(by_month,n,color=origin))+
  geom_line()+
  ggtitle("flights by month from each origin")
```

```{r}
dfc3=dfc2 |>
  group_by(origin,by_month) |> 
  summarise(passengers=sum(passengers))

dfc3 |>
  ggplot(aes(by_month,passengers,color=origin))+
  geom_line()+
  ggtitle("passengers by month from each origin")
```

Maybe an accumulated value would be better:

```{r}
dfc4=dfc3 |>
  mutate(cum_pass=cumsum(passengers))

dfc4 |>
  ggplot(aes(by_month,cum_pass,color=origin))+
  geom_line()+
  ggtitle("cummulative passengers from each origin") 

```

Let's look at the difference between those two curves:

For that computation, we will need to "pivot", making our dataframe wider.

```{r}
dfc5=dfc4 |> 
  pivot_wider(names_from = origin,
              values_from = c("passengers","cum_pass"))|>
  mutate(net_to_miami=passengers_LGA-passengers_MIA,
         cum_to_miami=cum_pass_LGA-cum_pass_MIA)
```

```{r}
dfc5 |>
  ggplot(aes(by_month,net_to_miami,color=origin))+
  geom_line()+
  ggtitle("monthly net flow to Miami") 
```

The data above is really monthly counts, not a continuous curve.   So - a better representation would be with bars.


```{r}
dfc5 |>
  ggplot(aes(by_month,net_to_miami,color=origin))+
  geom_col()
  ggtitle("monthly net flow to Miami") 
```

```{r}
dfc5 |>
  ggplot(aes(by_month,cum_to_miami,color=origin))+
  geom_line()+
  ggtitle("accumulated net flow to Miami") 
```




# Below here will not run
 
```{r eval=FALSE, include=FALSE}
df13=df11 |> mutate(percent_loading = passengers/seats*100)

df13 |> 
  ggplot(aes(percent_loading)) +
  geom_histogram()+facet_wrap(~carrier_name,scales="free_y")

```


### Summarize data

```{r eval=FALSE, include=FALSE}
df14=df2022 %>% filter(passengers>0,origin=="LGA" | dest=="LGA") %>% 
  group_by(month,carrier_name) %>% 
  summarise(total_pass=sum(passengers),  num_flights=sum(departures_performed)) %>%
  mutate(average_pass_per_flight=total_pass/num_flights)


df14 
```



